\documentclass[11pt]{article}

% Change "review" to "final" to generate the final (sometimes called camera-ready) version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
\usepackage[review]{acl}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{inconsolata}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\title{Anything2PBL: Hierarchical Constrained Generation with \\ Learner-Aware Scaffolding for Project-Based Learning}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  \texttt{email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  \texttt{email@domain} \\}

\begin{document}
\maketitle

\begin{abstract}
Project-Based Learning (PBL) is an effective approach for developing practical competencies through authentic project work.
However, creating high-quality training projects remains costly, as it requires not only engineering completeness but also learner-appropriate pedagogical design---in particular, deciding what constitutes a productive challenge versus tedious grunt work depends on the target learner rather than the project itself.
Existing approaches either focus on code generation without pedagogical consideration, or assist the learning process without producing a complete, runnable project.
We propose \textbf{Anything2PBL} (A2P), a hierarchical constrained generation framework that transforms arbitrary inputs into pedagogically-calibrated project repositories through a five-stage human-in-the-loop pipeline.
Experiments on a cross-disciplinary benchmark show that A2P consistently outperforms direct-prompting and chain-of-thought baselines, and that instructor calibration at the scaffolding stage yields disproportionately large gains.
\end{abstract}

\section{Introduction}

LLMs have made explanations cheap.
Given a paper, codebase, tutorial, or dataset, LLMs can easily produce summaries, slides, quizzes, and step-by-step guides.
Yet for many technical skills, the limiting factor is no longer access to explanations alone.
A learner does not master retrieval-augmented generation by reading about chunking strategies, embedding models, or hallucination mitigation; mastery requires making these design decisions inside a working system, observing failures, and repairing them.
What remains costly is therefore not the delivery of conceptual content, but the construction of practice environments in which learners must act on that content.

\noindent
We focus on a specific kind of practice environment: a runnable training repository tailored to a specified learner.
Unlike summaries, lesson plans, quizzes, or complete implementations, these repositories must satisfy two competing requirements.
They must run, so that learners receive authentic feedback from code, tests, data, and system behavior.
At the same time, they must leave the right parts unfinished, because these gaps are where learners perform the target cognitive and engineering work.
This is the artifact-construction setting of Project-Based Learning (PBL), where learning is organized around building authentic artifacts rather than only consuming instructional content~\cite{kilpatrick1918,thomas2000}.

\noindent
This makes PBL repository construction fundamentally different from ordinary software generation.
In software generation, an incomplete implementation is a defect to be repaired.
In a training repository, incompletion can be the teaching interface.
If the repository is too complete, the learner merely runs or lightly edits someone else's solution, inviting cognitive offloading~\cite{kelly2019cognitive,armitage2025offloading}.
If it is too incomplete, the learner faces overload rather than productive struggle~\cite{kapur2008productive,bjork2011desirable}.
If the wrong components are left unfinished, the task becomes prerequisite work, mechanical busywork, or unguided debugging.
The central problem is therefore not generation alone, but \emph{controlled incompletion}: determining what to implement, what to scaffold, what to leave to the learner, and how this allocation should vary with the learner profile.

\noindent
Controlled incompletion requires knowing what the project is supposed to teach.
A repository teaches RAG, for example, not because it contains RAG code, but because it makes the learner perform purposeful actions: choosing chunk boundaries, diagnosing retrieval failures, evaluating answer quality, and mitigating hallucinations.
The educational value of the repository is therefore not a property of the code alone.
It lies in a correspondence between the source material's concepts and dependencies, and the concrete actions the learner is required to perform.

\noindent
This correspondence cannot be recovered reliably from the final repository alone.
A generated project may run successfully while still assigning learners the wrong work, obscuring the intended learning path, or leaving gaps for reasons unrelated to pedagogy.
Direct prompting tends to collapse these design decisions into a single output artifact, making the pedagogical rationale difficult to inspect or revise.
Thus, PBL repository construction requires not only executable generation, but also an explicit representation of how source knowledge is transformed into learner-owned work.

\noindent
Existing LLM systems approach this problem from two neighboring directions.
Repository-level code generation systems optimize for executable completeness, using tests or benchmark task completion as the main signal of success~\cite{qian2023chatdev,hong2023metagpt,yang2024sweagent,wang2024openhands,jimenez2023swebench,li2024deveval}.
Educational content generation systems produce slides, lesson plans, textbooks, quizzes, or tutoring dialogue~\cite{zheng2025pptagent,sharma2025pustakai,zhang2024simclass,cohn2026edf}.
The former tends to eliminate implementation gaps; the latter rarely produces an executable environment in which learning objectives become learner-owned actions.
Neither line of work explicitly represents how source knowledge should become learner action, scaffolding, and validation in an executable project.

\noindent
We propose \textbf{Anything2PBL (A2P)}, a framework that turns arbitrary source artifacts into runnable, learner-calibrated PBL repositories.
A2P's central object is not the repository alone, but a \textbf{Pedagogical Repository Contract}: an inspectable intermediate representation that maps source concepts to learning objectives, learner-owned tasks, scaffolds, and validation checks.
The contract records what the learner should learn, where that learning is exercised in the repository, which parts are completed by the system or the learner, and how success is validated.
The repository is then generated as an implementation of the contract rather than produced as a single unconstrained generation.

\noindent
A2P constructs this contract through a five-stage human-in-the-loop pipeline: source analysis, project design, scaffolding calibration, repository construction, and validation.
The early stages identify what should be learned and how it can be embodied in a project; scaffolding calibration then decides which parts should be implemented by the system and which should remain learner-owned.
Repository construction and validation generate the runnable artifact while checking that these pedagogical decisions are preserved.
This turns incompletion from an accidental property of generated code into a design object that can be inspected, revised, and evaluated.

\noindent
We evaluate A2P on a cross-disciplinary benchmark spanning multiple source types and learner profiles.
Because the goal is not merely to generate runnable repositories, our evaluation considers both engineering validity and pedagogical validity.
We measure whether generated projects remain faithful to the source, embody concepts as learner actions, calibrate incompletion to the learner profile, avoid scaffold leakage, and preserve traceability from source concepts to validation checks.
A2P outperforms direct-prompting and chain-of-thought baselines, and ablations show that learner-aware scaffolding calibration is the dominant contributor to pedagogical quality.

\smallskip
\noindent In summary, our contributions are:
\begin{itemize}
    \item We formulate PBL repository construction as \textbf{pedagogical compilation}: transforming source artifacts into executable projects in which learning objectives become learner-owned actions.
    \item We introduce the \textbf{Pedagogical Repository Contract}, an explicit intermediate representation linking source concepts, learner tasks, scaffolding decisions, and validation checks.
    \item We implement \textbf{Anything2PBL}, a five-stage human-in-the-loop pipeline for generating runnable, learner-calibrated, and deliberately incomplete training repositories.
    \item We empirically show that A2P improves both engineering validity and pedagogical validity over direct generation baselines, with learner-aware scaffolding calibration contributing the largest gains.
\end{itemize}

\section{Related Work}

\subsection{Repository-Level Code Generation}

Multi-agent frameworks increasingly generate software at the repository level.
ChatDev~\cite{qian2023chatdev} and MetaGPT~\cite{hong2023metagpt} decompose development into role-based pipelines (analyst $\to$ coder $\to$ tester), while SWE-agent~\cite{yang2024sweagent} and OpenHands~\cite{wang2024openhands} equip agents with tool-augmented interfaces for navigating real codebases.
Self-collaboration code generation~\cite{dong2023selfcollaboration} demonstrates that role separation substantially improves Pass@1 on HumanEval and MBPP.
CodeAgent~\cite{zhang2024codeagent} further integrates five programming tools for repo-level synthesis.
Most recently, AutoP2C~\cite{shi2025autop2c} transforms academic papers into executable repositories through a four-stage pipeline---blueprint extraction, multimodal parsing, hierarchical decomposition, and iterative debugging---achieving full success where single-model baselines solve only 1 of 8 papers.
AutoP2C is architecturally closest to A2P: both perform ``input $\to$ executable repository'' transformation via hierarchical decomposition.
However, all these systems optimize a single objective---engineering correctness $E(O)$, validated by test pass rates~\cite{jimenez2023swebench,li2024deveval}---and none asks whether the generated repository is \emph{learnable}.
A2P differs by jointly optimizing $E(O)$ and pedagogical effectiveness $P(O)$: the same input yields different projects for different learner profiles.

\subsection{Educational Content Generation}

LLMs generate educational materials across diverse modalities.
Content-generation systems---PPTAgent~\cite{zheng2025pptagent}, DeepPresenter~\cite{zheng2026deeppresenter}, STORM~\cite{shao2024storm}, and PustakAI~\cite{sharma2025pustakai}---produce static artifacts such as slides, articles, and textbooks.
LessonPlanner~\cite{wang2024lessonplanner} grounds lesson plan generation in Gagn\'{e}'s nine events of instruction.
Tutoring systems take a complementary approach: SimClass~\cite{zhang2024simclass} simulates classroom interactions with multi-agent role-play, MedTutor~\cite{jang2026medtutor} provides case-based medical education, and EDF~\cite{cohn2026edf} implements theory-driven adaptive scaffolding.
Co-design studies with K-12 educators identify PBL project design as the primary challenge where LLM assistance is most requested~\cite{lee2025codesigning}.
However, all existing systems produce a single modality---slides, lesson plans, dialogue turns, or articles---rather than a complete, runnable project repository.
The gap between generating educational \emph{text} and generating an executable \emph{learning artifact} remains open; A2P targets precisely this gap.

\subsection{Scaffolding and Learner Modeling}

A2P's learner-aware calibration draws on two theoretical traditions.
Vygotsky's zone of proximal development~\cite{vygotsky1978} and Wood et al.'s scaffolding framework~\cite{wood1976scaffolding} establish that effective instruction targets the region between independent capability and assisted capability.
Sweller's cognitive load theory~\cite{sweller1988cognitive} provides the complementary lower bound: when working memory is overloaded, learning fails regardless of task design.
Puntambekar and H\"{u}bscher~\cite{puntambekar2005scaffolding} extend scaffolding to technology-mediated environments, identifying three critical features---ongoing diagnosis, calibrated support, and fading---that A2P operationalizes through its learner profile mechanism.
The tension between minimal guidance and structured support is well-documented: Kirschner et al.~\cite{kirschner2006minimal} argue that pure discovery learning imposes excessive cognitive load on novices, while Hmelo-Silver et al.~\cite{hmelosilver2007scaffolding} respond that effective PBL is heavily scaffolded inquiry, not minimal guidance.
A2P resolves this tension by calibrating scaffold density per learner profile, following Krajcik and Shin's~\cite{krajcik2014pbl} characterization of PBL as requiring both authentic artifacts and structured support.

On the computational side, learner modeling has evolved from Bayesian Knowledge Tracing~\cite{corbett1995knowledge} through deep knowledge tracing~\cite{piech2015deep} to LLM-augmented approaches such as SINKT~\cite{gao2024sinkt}, which uses LLM-encoded concept semantics for inductive knowledge tracing.
These systems model learners for \emph{runtime} adaptation---selecting or sequencing existing content.
A2P instead uses a one-shot learner profile to condition \emph{generation}, deciding what to scaffold, what to leave as productive struggle~\cite{kapur2008productive}, and what to omit as noise \emph{before} the project exists.

\subsection{Hierarchical Planning and Human-in-the-Loop Generation}

A2P's five-stage pipeline draws on hierarchical task decomposition.
Plan-and-Solve prompting~\cite{wang2023planandsolve} shows that planning before execution improves zero-shot reasoning; Least-to-Most prompting~\cite{zhou2023leasttomost} demonstrates that sequential sub-problem solving enables compositional generalization; and Hi-CoT~\cite{zhang2026hicot} alternates between instruction planning and step execution, reducing reasoning trace length while improving accuracy.
In multi-agent settings, self-collaboration pipelines~\cite{dong2023selfcollaboration} confirm that role-specialized stages outperform monolithic generation.
A2P applies this principle concretely: its first two stages (topic analysis and project design) correspond to the planning phase, while the latter three (scaffolding, code generation, and validation) execute the plan with stage-specific constraints.

The human-in-the-loop dimension is equally critical.
Amershi et al.'s~\cite{amershi2019guidelines} 18 guidelines for human-AI interaction organize intervention points along a temporal axis.
Empirical analysis of writing collaboration reveals that a small set of prototypical behaviors---revising intents and injecting content---dominate human-AI interaction~\cite{mysore2025paths}, and interaction-required interfaces promote ownership and cognitive engagement over passive acceptance~\cite{lee2025interaction}.
A2P's five intervention points constitute deep interaction: the instructor reviews and revises at each stage boundary rather than accepting a single end-to-end output.
This design reflects the finding that HITL workflows are necessary under quality-assurance constraints~\cite{wang2025summit} similar to those in educational content generation.

\bibliography{custom}

\end{document}
