\section{Introduction}

Curating assignments is a crucial step in education, especially when the aim of learning is to produce practical skills.
A well orchestrated assignment can help learner to ... \nn{Add citations showing learn by do is important and helpful.}
However, crafting such assignments is a non-trivial task and takes substantial times of teachers as it requires the joint design of aaa, bbb, ccc.
Thus it is necessary to explore the potential of automating the designing process of learning assignments.


%High-quality training projects are essential for practical technical education.
% Many skills are learned not from instruction alone, but through sustained work on realistic tasks.
% However, high-quality training projects remain difficult to create.
% A useful project must contain enough structure for learners to make progress, while leaving enough work for them to practice the target skills.
% This balance depends on the learner: the same task can be productive for one learner but too easy, tedious, or overwhelming for another.
% As a result, high-quality personalized training projects remain scarce and costly to produce.

As shown in Figure~\ref{fig:demo}, designing a learning assignment involves the crafting of multiple materials:
1. \textbf{Project Materials.} Students should be given with a set of project materials such as start up files, instructions, requirements. These materials provides initial information
2. \textbf{Instructional Procedure.}


Producing such projects is a multi-facit challenge:
It has to be a carefully designed training repository, while remaining incomplete in the places where the learner should practice.
Those places depend on the learner's background, goals, and expected level of support.
Creating such a project is both an engineering and a pedagogical problem: the system must decide what to build, what to scaffold, and what to leave for the learner.
These decisions interact: changing the project goal can change the required scaffolding, and changing the scaffolding can change what counts as meaningful learner work.


Existing LLM systems cover important pieces of this process, but not the full problem.
Repository-level agents generate and modify runnable code-bases~\cite{qian2023chatdev,hong2023metagpt,yang2024sweagent,wang2024openhands,zhang2024codeagent,shi2025autop2c}. But these systems are built to complete software, whereas a training project must preserve parts of that work for the learner to do.
On the other hand, educational systems generate slides, QA, textbooks, and classroom interactions~\cite{zheng2025pptagent,zheng2026deeppresenter,shao2024storm,sharma2025pustakai,wang2024lessonplanner,zhang2024simclass}; these produce instructional content rather than an executable environment in which learners build and practice.
Tutoring, scaffolding, and learner-modeling systems adapt support during learning~\cite{jang2026medtutor,cohn2026edf,deriyeva2026script,gao2024sinkt}, but typically assume that the task or material already exists.
Together, these gaps leave open a problem: how to turn a learner's goals into a concrete project design, and then realize that design as a runnable repository with appropriate learner work.


We propose \model, a framework that transforms learner intent into runnable training repositories tailored to individual learners.
\model interactively refines the learner's goals into a concrete project blueprint.
The blueprint specifies what the project will ask the learner to build, which skills it should exercise, what support should be provided, and what work should remain for the learner.
The system then expands this blueprint into a runnable training repository, including tasks, scaffolds, files, tests, and validation behavior.
To keep the blueprint aligned with the learner, \model asks clarifying questions when key design choices are uncertain, rather than hiding those choices inside the generated repository.



\smallskip
\noindent In summary, our contributions are:
\begin{itemize}
    \item We formulate PBL repository construction as \textbf{pedagogical compilation}: transforming source artifacts into executable projects in which learning objectives become learner-owned actions.
    \item We introduce the \textbf{Pedagogical Repository Contract}, an explicit intermediate representation linking source concepts, learner tasks, scaffolding decisions, and validation checks.
    \item We implement \textbf{Anything2PBL}, a five-stage human-in-the-loop pipeline for generating runnable, learner-calibrated, and deliberately incomplete training repositories.
    \item We empirically show that A2P improves both engineering validity and pedagogical validity over direct generation baselines, with learner-aware scaffolding calibration contributing the largest gains.
\end{itemize}